### 2014-10-24 thesis/method/selection.tex
Finished preprocessing section, restructure of overall methods section

\chapter{Feature Preprocessing and Selection} \label{chap:features}

In the previous chapter we began by describing each training example as a
vector of feature values, and then described how a range of different
learning algorithms used those feature vectors to learn a way to
distinguish examples of one class from another. Ideally, all features given
in a data set should be relevant and useful in not only describing the training
example but also predicting the class.
In practice, and particularly
in medicine, a large number of attributes about a patient is usually recorded.
This is detrimental for two reasons: the training time is significantly
increased because the classifier must incorporate irrelevant information, and
this irrelevant information tends to degrade the performance
of even state-of-the-art classifiers such as C4.5 decision trees
\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
of the most relevant features; to find them, we will use feature selection.

It is also common for data to require preprocessing before a classifier can
be trained upon it -- indeed, even before feature selection is performed.
Otherwise, it will be hard to know if the results we see are reflective of
the algorithms we use or simply a case of lucky data. Preprocessing involves
cleaning and transforming the data, and is critical to the performance of
learning algorithms that will be used to learn from it.

In this chapter we describe the various preprocessing steps performed on
the data, as well as the feature selection methods we evaluate in our work.

\section{Preprocessing}
In any data mining task, adequately preparing the data ensures the best
chance of success: this includes deciding how to deal with missing and
inaccurate values, as well as appropriately transforming the data
\citep{Witten2005} and selecting or constructing the most suitable features
for the task \citep{Kotsiantis2006}.
In this section we describe how we addressed each of
the major issues and the reasons for our choice.

\subsection{Cleaning} % missing values, irrelevant features
The exact steps involved in \textit{cleaning} the data will vary between data
sets, but the overall process is similar. Common issues to deal with include:
\begin{itemize}
\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
data entry
\item Leading or trailing whitespace, and other garbage characters,
perhaps as a result of data entry or
the system that stores the data
\item Missing values for various features or for the class label
\item Values for features that do not make sense (such as a patient age of
-1)
\end{itemize}

Since the data was initially in an Excel spreadsheet, we explored the data
by hand using Microsoft Excel's filter function on each column and
consolidated many categorical values simply
by correcting spelling and punctuation. For example:
\begin{itemize}
  \item In the \textbf{sex} attribute, there were values of `Female' and
  `Femal' which were consolidated into `Female'.
  \item In the \textbf{mechanism} attribute, which was textual and indicated
  the category of how the patient was injured, there were such categories as
  `Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
  Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
  respectively.
  \item Some categories in \textbf{mechanism} were also divided into whether or
  not the mechanism of injury was self-inflicted, such as `Shooting' and
  `Shooting-selfinflicted'. When we asked the domain expert, Michael Dinh, his
  suggestion was to combine them into one category: in this case, simply
  `Shooting'; this was done for all the categories in this attribute.
  \item \textbf{disp from ED} indicated the section of the hospital that the
  patient was discharged to. This was also a textual categorical attribute.
  There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
  all combined into one category, `GICU'.
\end{itemize}

The data was then exported into CSV (comma-separated value) format, so that
we could continue to process it programmatically rather than manually. Many
fields in the data set contained leading or trailing whitespace, which was
removed with a few lines of Python code (see appendix \todo{[add reference
to appendix]}). All attribute names were converted to lowercase. In the case
where attribute names consisted of several words, the spaces were replaced
with underscores (\_) to ensure that the full attribute name was read.
For example, ``Any Cancer'' was converted into ``any\_cancer''.

In order for our results to be comparable to that of Dinh
et. al. \citep{Dinh2013a}, we removed all instances with missing values,
leaving 2517 instances with which to use to create our prediction models.

\subsection{Normalisation}
All numeric attributes were normalised into the range $[-1,+1]$

\subsection{Discretisation}

\section{Feature Selection}
After data preparation, selecting a subset of features or attributes to use
in a model is also important, and helps to reduce noise and redundant
attributes from becoming part of the final model. As we mentioned earlier,
experiments have found that
the performance of various classification methods such as decision trees
\textit{deteriorates} in the presence of useless attributes.
Our data set contains 73
non-target attributes, which is a fairly high number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \textit{age65} feature) is directly dependent on the
value of their \textit{age} attribute.

>>>>

\chapter{Feature Selection} \label{chap:selection}

Recall that each feature tells
us something about the training example. Ideally, it should also help us
classify the example. However, in practice and particularly
in medicine, a large number of attributes about a patient is usually recorded,
not all of which are relevant in predicting the class.
Having such a large number of features
is detrimental for two reasons: the training time is significantly
increased because the classifier must incorporate irrelevant information, and
this irrelevant information tends to degrade the performance
of even state-of-the-art classifiers such as C4.5 decision trees
\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
of the most relevant features; to find them, we will use feature selection.

\section{Feature Selection}
After data preparation, selecting a subset of features or attributes to use
in a model is also important \citep{Kotsiantis2006},
and helps to reduce noise and redundant
attributes from becoming part of the final model. As we mentioned earlier,
experiments have found that
the performance of various classification methods such as decision trees
\textit{deteriorates} in the presence of useless attributes.
Our data set contains 73
non-target attributes, which is a fairly high number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \textit{age65} feature) is directly dependent on the
value of their \textit{age} attribute.
